### 2024-08-15 paper/main.tex
at kleene proof

\documentclass[12pt,letterpaper]{article} 
\usepackage[letterpaper,margin=1in]{geometry} 
\usepackage{mathptmx}
\usepackage[singlespacing]{setspace}
\usepackage{fancyhdr}
\setlength{\headheight}{14.49998pt}
\usepackage{relsize}
\usepackage[bottom]{footmisc}
\usepackage{tabularx}
\setlength{\parindent}{2em}
\usepackage{url}
\usepackage{hyperref}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage{expex}
\usepackage{amsfonts} 
\usepackage{tikz}
\bibpunct[:]{(}{)}{;}{a}{}{,}
\usepackage{bibentry}
\usepackage{acronym}
\usepackage{multicol}
\author{Ruben Triwari}
\usepackage{scrhack} 
\usepackage{microtype}
\usepackage[acronym,xindy,toc]{glossaries}
\makeglossaries
\loadglsentries{glossary.tex}
\usepackage{ragged2e}
\babelhyphenation[english]{
an-oth-er
ex-am-ple
}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}[section]
\newenvironment{proof}{\paragraph{Proof:}}{\hfill$\square$}
\usepackage{dcolumn}
\newcolumntype{2}{D{.}{}{2.0}}
\usetikzlibrary{
  automata, positioning,
  arrows, matrix, decorations.pathreplacing,
  shapes.geometric
}
\newcommand{\fps}[1] {
\mathbb{#1}\langle \langle \Sigma^* \rangle \rangle
}

\tikzset{ 
  table/.style={
    matrix of math nodes,
    row sep=-\pgflinewidth,
    column sep=-\pgflinewidth,
    nodes={rectangle,text width=3em,align=center},
    text depth=1.25ex,
    text height=2.5ex,
    nodes in empty cells,
    left delimiter=[,
    right delimiter={]},
    ampersand replacement=\&
  }
}

\begin{document}

\begin{center}\uppercase{Ludwig-Maximilians-Universität München}\end{center}
\begin{center}\uppercase{Chair of Theoretical Computer Science and Theorem Proving}\end{center}

\vspace*{10mm}
\begin{center}
\includegraphics[height=40mm]{sigillum.png}
\end{center}
\vspace*{10mm}

\title{Formal Languages and Automatas}
\date{\vspace{-3ex}}
{\let\newpage\relax\maketitle}
\thispagestyle{empty}

\begin{center}
\begin{large}
\begin{Large}
Seminar Paper on Formal Languages and Automatas\\
\end{Large}
for the course ``Algebra \& Computer Science'' \\
\end{large}
\end{center}
\vspace{1cm}
\begin{center}
\begin{large}
Supervisor: Prof. Dr. Jasmin Blanchette\\
\end{large}
\end{center}
\begin{center}
\begin{large}
Advisor: Xavier Genereux\\
\end{large}
\end{center}

\begin{center}
\begin{large}
Submission Date: \date{\today} \\
\end{large}
\end{center}

\vspace{1,5cm}
\cleardoublepage{}
\thispagestyle{empty}
\vspace*{0.8\textheight}
\noindent
\makeatletter
\begin{center}
{\normalfont\bfseries} Disclaimer
\end{center}
\begin{flushleft}
I confirm that this seminar paper is my own work and I have documented all sources and material used.\\
\makeatother
\vspace{15mm}
\noindent
Munich, \date{\today} \hspace{50mm} Author
\end{flushleft}
\cleardoublepage{}

\section*{Abstract}
Abstract...
\textbf{Keywords}: Math

\newpage
\tableofcontents
\newpage

\setcounter{page}{1}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[R]{\thepage}
\renewcommand{\headrulewidth}{0pt} %obere Trennlinie

\section{Introduction}
\section{Backround}
First of all, we need to define some structures and ideas,
that will be really important in this paper. We will define 
formal languages in two ways and then automatas arcordingly.
A formal language is just a set of words, where words are 
a series of characters from an alphabet.
\begin{definition}
  An Alphabet is an abritrary non empty finite set denoted with $\Sigma$
\end{definition}
\begin{example} Simple examples for Alphabets would be:
  \begin{enumerate}
    \item $\Sigma = \{x\}$
    \item $\Sigma = \{a,b,c,d, \dots z \}$
    \item $\Sigma = \{1,2,3, \dots n \}$ where $n \in \mathbb{N}$
  \end{enumerate}
\end{example}
This definition is rather general, because an alphabet $\Sigma$
can really be anything. We could also define an Alphabet as subset of
the natural numbers $\Sigma \subset \mathbb{N}$. This is equalivant to
the previous definition, becaus we can always find a one-to-one map between
a finite amount of objects and the natural numbers. In other words 
we can always enumerate a finite amount of objects. The above definition
results in more readable words, thus resulting in more readable problems.
Furhtermore its unintuitive to have words full of numbers and languages
full of number series. That is why im chosing this definition, like most
of the literature.\\
Now with our first building block defined we can define words, which are just
a series of characters in an alphabet.
\begin{definition}
  Let $\Sigma$ be an fixed alphabet, then $w$ is a series of characters out of
  $\Sigma$. 
  \[
    w = (a_1,a_2, \dots, a_n) \in \Sigma^n
  \]
  We will denote it by:
  \[
    w = a_1a_2\dots a_n
  \]
  The empty word which contains no characters, 
  and is denoted by $\epsilon$.
  %book schützenbert
\end{definition}
\begin{example}
  Simple examples for words would be:
  \begin{enumerate}
    \item $\Sigma = \{a,b,c\} \rightsquigarrow w_1 = abc$
    \item $\Sigma = \{x\} \rightsquigarrow w_2 = xxxx$
    \item $\Sigma = \{1,2,3\} \rightsquigarrow w_3 = 112233$
    \item $\Sigma = \{1,2,3\} \rightsquigarrow w_4 = \epsilon$
  \end{enumerate}
\end{example}
The empty word $\epsilon$ is analogous to the empty set in Set theory.
It has a lot uf uses, but one obvious is that now one can define 
a Monoid over words $(M, \cdot, \Sigma, \epsilon)$.
M is a set of words and multiplication is concetanation of words, then 
$\epsilon$ is the neutral element of the Monoid. Which brings us to
the next definition.
\begin{definition}
  Let $\Sigma$ be an fixed alphabet and  $w$, $v$ words with 
  characters out of $\Sigma$. Then the contenation is defined by:
  \[ w \cdot v := wv\]
  sometimes just written as $wv$. With powers recursively defined by:
  \begin{align*}
    w^0 &:= \epsilon\\
    w^1 &:= w\\
    w^{n+1} &:= w^n\cdot w
  \end{align*}
\end{definition}
In simple terms a concatenation of two words is just a new word with 
the two words chained togehter.\\
Now we can again build on words to define langauges.
\begin{definition}
  Let $\Sigma$ be an fixed alphabet, then a formal language is a 
  set of words, with characters in $\Sigma$.
\end{definition}
\begin{example}
  Simple examples for formal languages would be:
  \begin{enumerate}
    \item $\Sigma = \{a,b,c\} \rightsquigarrow L = \{aaa, bbb, ccc, abc\}$
    \item $\Sigma = \{x\} \rightsquigarrow L = \{ x, xx, xxx, xxxx\}$
    \item $\Sigma = \{1,2,3\} \rightsquigarrow L = \{11, 22, 33, 123,\epsilon\}$
  \end{enumerate}
\end{example}
This definition is really clear and simple. We will see a less intuitive way to
define language shortly. Obviously the simpler one has some downsides, 
which we will discuss later in the paper. \\
The kleene star $^*$ is a really powerful operator, to define it we need
to first define contenation for languages.
\begin{definition}
  Let $\Sigma$ be an fixed alphabet and  $L$, $V$ languages with 
  words constructed out of $\Sigma$.
  Then the contenation of formal languages is defined by:
  \[ L \cdot V := \{ l \cdot v | l \in L \land v \in V \}\]
  sometimes just written as $LV$. With powers defined recursively:
  \begin{align*}
      L^0 & := \{\epsilon \}\\
      L^1 & := L  \\
      L^{i+1} & := \{ l_i\cdot l | l_i \in L^i \land l \in L\}
  \end{align*}
  % wikipedia kleene star
\end{definition}
In simple terms we are creating a new language by compbining every word
of the first language with every word of the second.
Now we are able to define the kleene star operator.
\begin{definition}
  Let $\Sigma$ be an fixed alphabet and  $L$  a language with 
  words constructed out of $\Sigma$.
  Then the kleene star of  a formal language is defined by:
  \[ 
    L^* := \bigcup_{i \in \mathbb{N}_0} L^i = 
    L^0 \cup L^1 \cup L^2 \cup \dots
  \]
\end{definition}
In simple terms it generates all possible permutation of abriatary length
of a formal language. This is really powerful, because $\Sigma^*$ is the
biggest language one can construct from the alphabet $\Sigma$. Thus every 
language $L$ generated by $\Sigma$ is just a subset of $\Sigma^*$.\\
Unintuively, we can also write a formal language as a formal power series.
\begin{definition}
  Let $\Sigma$ be an fixed alphabet, then a formal language can be 
  written as formal power series $L$:
  \[
    L: \Sigma^* \to \mathbb{B}
  \]
  where
  \[
    \mathbb{B} = \{0, 1\}
  \]
  \[
    L = \sum_{w \in \Sigma^*} (L,w)w \verb| where | 
    (L,w) \in \mathbb{B}.
  \]
  {\bf Note:} The set of formal power series is denoted by 
  $\mathbb{B}\langle \langle \Sigma^* \rangle \rangle 
    = \{ \Sigma^* \to \mathbb{B}\}$\\
  {\bf Notation:} If the language consist only out of one 
  character or the empty word $a \in \Sigma \cup \{\epsilon\}$, 
  then the formal power series would be: 
  \[
    L = h \in \fps{B}
  \]
  where 
  \[
    h: \Sigma^* \to \mathbb{B}
  \]
  \[
    h(w) = 
    \begin{cases}
      1 & \text{if   } w=a \\
      0 & \text{otherwise}
    \end{cases}.
  \]
  In this case we just write $L = a$.
    %schützenberg
\end{definition}
This looks like a sum, we call it series, but it is actually a map from 
words to zero or one. The sum is a way to write a function definition: Infront
of every word $w \in \Sigma^*$ in the series, there is a $x \in \mathbb {B}$,
which is zero or one. So we can interpret every element in the series as an 
assignment: $w \mapsto x$.
All words that map to one, are in the language, the others are not. So we can write a formal language
as a function $\Sigma^* \to \mathbb{B}$.\\
The representation with sets has the andvantage that a formal language
will inherit all operations on sets like complement, union, and intersection.
Those operations are really useful, thus we will define similar on the power
series version.
\begin{definition}
  Let $\Sigma$ be an fixed alphabet and $L, V \in \mathbb{B}\langle \langle
  \Sigma^* \rangle \rangle$, where
  \[
    L = \sum_{w \in \Sigma^*} (L,w)w \hspace{0.5cm} \text{and} \hspace{0.5cm}
    V = \sum_{w \in \Sigma^*} (V,w)w.
  \]
  Addition and multiplication on formal power series is defined
  as follows:\\
  \[
    L + V := \sum_{w \in \Sigma^*} ((L,w) + (V,w))w
    \hspace{0.5cm} \text{and} \hspace{0.5cm} 
    L \cdot V := \sum_{\substack{u,v \in \Sigma^* \\ w = uv }} 
      ((L,u) \cdot (V,v))w
  \]
  This definition is not recursive, because the right addition and 
  muliplication operater are from the Boolean Algebra 
  $\mathbb{B} = \{0,1\}$.
  \begin{center}
    \begin{tikzpicture}
      \matrix (m) [nodes={minimum width=2em,minimum height=2ex},matrix of nodes]
      {
      $+$&0&1\\
      0&0&1\\
      1&1&1\\
      };
      \draw[very thick] (m-1-1.north east) -- (m-3-1.south east);
      \draw[very thick] (m-1-1.south west) -- (m-1-3.south east);
      \draw (m-1-1.north west) -- (m-3-1.south west);
      \draw (m-1-1.north west) -- (m-1-3.north east);
      \foreach \x in {2,...,3}{
        \draw (m-1-\x.north east) -- (m-3-\x.south east);
        \draw (m-\x-1.south west) -- (m-\x-3.south east);
      }

      \matrix [
        nodes={minimum width=2em,minimum height=2ex}
        ,matrix of nodes
      ] at (4,0)
      {
        \node(a){$\times$}; & \node(b){0}; & \node(c){1}; \\
        \node(d){0}; & \node(e){0}; & \node(f){0}; \\
        \node(g){1}; & \node(h){0}; & \node(i){1};\\
      };
      \draw[very thick] (a.south west) -- (c.south east);
      \draw[very thick] (a.north east) -- (g.south east);

      \draw (a.north west) -- (c.north east);
      \draw (a.north west) -- (g.south west);

      \draw (b.north east) -- (h.south east);
      \draw (d.south west) -- (f.south east);

      \draw (c.north east) -- (i.south east);
      \draw (g.south west) -- (i.south east);

    \end{tikzpicture}
  \end{center}
  With powers defined recursively:
  \begin{align*}
    L^0 &:= \epsilon\\
    L^1 &:= L\\
    L^{i+1} &:= \sum_{\substack{u,v \in \Sigma^* \\ w = uv}}
    ((L^i,u)\cdot (L,v))w, 
  \end{align*}
  Kleene Star can now also be defined for formal power series:
  \[
    L^* := L^0 + L^1 + L^2 + L^3 + \dots = \sum_{i \in \mathbb{N}_0} L^i.
  \]
\end{definition}
With these operation defined our two versions of formal languages are 
powerful tools, which will be compared in the following sections. In order
to conpare them we will also need finite automatas. Simillar to formal languages
there are multiple ways one can define finite automatas. In this paper we
will define two. One is more suitable for combining it with the set version
of formal languages and the other one is more siutable for the formal power
series representation. First we will define the one more suitable for the
set version.
\begin{definition}
  A finite automata is a quintuple $M = (Q, \Sigma, \delta, q_0, F)$, 
  where
  \begin{enumerate}
    \item Q is a finite set of states,
    \item $\Sigma$ is an alphabet, all inputs are constructed from,
    \item $\delta: Q \times \Sigma \to Q$ is the transition function,
      carrying the automata from one state to another,
    \item $q_0 \in Q$ is the initial state of the automata,
    \item $F \subset Q$ are the accepting states of the automata.
  \end{enumerate}
  %Definition is out of theoretical computer science book (german)
\end{definition}
In simple terms, an finite automata gets an input string and a starting 
state. Then it works through the input string, starting at the first charater,
carrying the starting state to another according to the transition function.
Exactly the same procedure, for the second input string character and the 
remainding characters. If this process ends up in an accepting state, 
the automata accepts the input string. Consequently, 
the finite automata generates a language, namely the words that the automata 
accepts. The definition above is one possible way to encode a finite automata. 
Like with graphs we can encode the transition function of an finite
automata, with just a matrix.
\begin{definition}
  Let $M = (Q, \Sigma, \delta, q_0, F)$ be an abitrary finite 
  automata, where $Q = \{ s_1, s_2, \dots s_n\}.$ \\
  Frthermore, we need a helper function:
  \[
    \delta': \mathbb{N} \times \mathbb{N} \to \fps{B}
  \]
  where
  \[
    \delta'(i, j) = 
      \begin{cases}
        \alpha, & \text{if} \hspace{0.2cm} \exists \alpha \in \Sigma: 
          \delta(s_i, \alpha) = s_j \\
        0, & \text{otherwise}
      \end{cases}.
  \]
  Then the finite automata $M$ can be encoded with a Matrix
  $A \in \fps{B}^{n \times n}$
  in the following way:
  \[
    A = 
    \begin{pmatrix}
      \delta'(1,1)       & \delta'(1,2) & \delta'(1,3) & \dots & \delta'(1,n) \\
      \delta'(2,1)       & \delta'(2,2) & \delta'(2,3) & \dots & \delta'(2,n) \\
      \vdots             & \vdots       & \vdots       & \ddots& \vdots\\
      \delta'(n,1)       & \delta'(n,2) & \delta'(n,3) & \dots & \delta'(n,n) \\
    \end{pmatrix}
  .\]
  We can denote a state transition $s_i \to s_j$ with $s_iA_{ij} = s_j$.\\
  {\bf Note:} We definined $A \in \fps{B}^{n \times n}$ and not 
  $A \in \Sigma^{n \times n}$, because otherwise the defintion is not closed
  under matrix multiplication, nor matrix addition.\\
  The kleene star of a Matrix can intuively  be defined as:
  \[
    A^* = E_n + A + A^2 + A^3 \dots  = \sum_{i \in \mathbb{N}_0} A^i
  \]
  where
  \[ E_n \in \fps{B}^{n \times n}\]
  \[
    (E_n)_{ij} := 
    \begin{cases}
      \epsilon & \text{if } \hspace{0.2cm} i=j \\
      0        & \text{otherwise}
    \end{cases} \in \fps{B}
  \]
  Now we can write our finite automata in a shorter way:
  \[
    M' = (\Sigma, A, q_0, F)
  \]
\end{definition}
This definition only works, because we have finite sets. Note that we can omit
the state set, since as mentioned above, there always exist a one-to-one map
between a set with finite objects and a subset of the natural numbers. The first
object gets the number one, the second the number two, and so forth. The new
definition of the state transition function works like a lookup table, when one
wants to know what characters map state $i$ to state $j$, its a simple lookup:
$M_{ij}$.

\begin{example}
  Here is a simple automata depicted, where the circles are states
  , the arrows are transitions, the start is indicating the initial 
  state, and the double lined circle is the accepting state.
  \[ 
    M_\text{tuple} = (\{s_1, s_2\}, \{a,b,c\}, \delta_0, s_1, \{s_2\}),
    M_\text{matrix} = (\{a,b,c\}, A, s_1, \{ s_2 \})
  \]\\
  where
  \[
    \delta_0(s_1, a) = s_1, \hspace{0.2cm} 
    \delta_0(s_1, b) = s_2, \hspace{0.2cm} \delta_0(s_2,c) = s_2
  \]
  and
  \[ 
    A = 
    \begin{pmatrix}
      a & b \\
      0 & c 
    \end{pmatrix} \in \fps{B}^{2\times 2}.
  \]
  \begin{center}
    \begin{tikzpicture}[shorten >=1pt,node distance=2cm,on grid,auto]
      \node[state,initial] (q1) at (0,0) {$s_1$};
      \node[state,accepting] (q2) at (3,0) {$s_2$};
      \path[->] (q1) edge[loop above] node{a} (q1)
      (q2) edge[loop above] node{c} (q2)
      (q1) edge[bend left] node{b} (q2);
    \end{tikzpicture}
  \end{center}
  The resulting accepting formal language is:
  \[ \text{Set represntation: } L = \{a^nbc^n : n \in \mathbb{N}\}\]
  \[
    \text{Formal power series represntation: } L = a^*\cdot b\cdot c^* 
  = a^*bc^*
\]
\end{example}
% Define Formal languages (Sets and Power sereies) 
% and automatas (Tuple and Matrix)
\section{Formal Languages: Sets vs. Formal Power Series}
In this chapter we will discuss the obvious diffrences and 
simillarities of both representations. First we will give intuition,
why one might use a seires to represent a function definition.
Secondly, we will map the formal power series version into the set
version and the other way around, to get a better understanding of both.
Lastly, we will discuss why formal power series are more powerful
than just sets.\\
The intuition behind the notation of formal power series is quite simple.
The operation defined on the formal power series, which are maps, 
are really close to a series. For addition, formal power series and 
normal series are quite simillar, namely defined pointwise.\\
In the following are $L,V \in \fps{B}$ and $A,B$ series.
\[ 
  L + V = \sum_{w \in \Sigma^*} ((L,w) + (V,w))w \hspace{0.3cm}
  \leftrightarrow \hspace{0.3cm}
  A + B = \sum_{n \in \mathbb{N}} (a_n + b_n) x^n
\]
The multiplication defined on formal power series are again really close
to the cauchy product of series.
In the following are $L,V \in \fps{B}$ and $A,B$ series.
\[
  L \cdot V = \sum_{\substack{u,v \in \Sigma^* \\ w=uv}} 
  ((L,u) \cdot (V,v))w
  \hspace{0.3cm} \leftrightarrow \hspace{0.3cm}
  A \cdot B = 
  \sum_{\substack{i,j \in \mathbb{N} \\ n = i + j}} (a_i \cdot b_j) x^n
\]
Consequently, it is quite handy to use a formal power series as function 
definition, because the operations seem natural with series.
Despite the notation being quite confusing at first glance.\\
%Maybe subsection
Next we will show that bot representations of a formal language are 
equivalent, by mapping one to another.
To be sure that both representation are actually equivalent, we need to
show there exist an bijective map, between formal power series and the
set of words. This map needs to preserve the structure, that means the
order of function application and operator application must be irrelevant.
Such maps are called isomporphism and structures for that a isomorphism 
exist are called isomporphic to each other. Thus we want tho show that
both representation are isomorphic.
\begin{proof}
  Let $\Sigma$ be an fixed alphabet, then we claim that:
  \[
    \mathbb{B}\langle \langle \Sigma^* \rangle \rangle
    \cong \mathcal{P}(\Sigma^*)\]
  In order to show those are isomorphic, we provide a map $\phi: 
  \mathbb{B}\langle \langle \Sigma^* \rangle \rangle \to \mathcal{P}(\Sigma^*)$ 
  and its inverse $\phi^{-1}: \mathcal{P}(\Sigma^*) \to \fps{B}$ and
  show that $\phi$ is bijective and preservers structure, formally:\\
  Let $L,V \in \fps{B}$, then for $\phi$ must hold:
  \begin{align}
    \phi(L + V) &= \phi(L) \cup \phi(V) \\
    \phi(L \cdot V) &= \phi(L) \cdot \phi(V)\\
    \phi(L^*) &= \phi(L)^*
  \end{align}
  We define $\phi$ and its inverse as follows:
  \[ 
    \phi(L) = \{ w \in \Sigma^* : (L,w) = 1\}
  \]
  \[
    \phi(L)^{-1} = h_{L}
  \]
  where
  \[ h_L: \Sigma^* \to \fps{B}\]
  \[
    h_L(w) = 
    \begin{cases}
      1, & \text{if} \hspace{0.2cm} w \in L \\
      0, & \text{otherwise}
    \end{cases}
  \]
  To prove that $\phi$ is bijective, it must hold
    \[ 
      \phi^{-1}(\phi(L)) = L
    \]
    and 
    \[
      \phi(\phi^{-1}(L)) = L.
    \]
    First we show $\phi^{-1}(\phi(L)) = L$\\
    Let $\ L \in \fps{B}$ abitrary but fixed, with 
    \[ L =\sum_{w \in \Sigma^*} (L,w)w. \]
    Then
    \begin{align*}
      \phi^{-1}(\phi(L)) = \phi^{-1}(\{ w \in \Sigma^* : (L,w) =1 \})
       &= 
       \begin{cases}
        1 & \text{if} \hspace{0.2cm} w \in \{ w \in \Sigma^* : (L,w) =1 \}\\
        0 & \text{otherwise}
       \end{cases}\\
       &= 
       \begin{cases}
        1 & \text{if} \hspace{0.2cm} (L,w) = 1\\
        0 & \text{otherwise}
       \end{cases}\\
       &= 
       \sum_{w \in \Sigma^*} (L,w)w = L.
    \end{align*}
    Next we need to show that $\phi(\phi^{-1}(L)) = L.$\\
    Let $\ L \in \mathcal{P}(\Sigma^*)$ abitrary but fixed, then
    \begin{align*}
      \phi(\phi^{-1}(L)) = \phi(h_L) = \{w \in \Sigma^* : (h_L,w) = 1 \}
      = \{ w \in \Sigma^* : h_L(w) = 1\} = \{ w \in \Sigma^* : w \in L\} = L.
    \end{align*}
    Consequently, $\phi$ is a one-to-one correspondence between formal power
    series and set of words.
    We are not done yet, we still need to prove that $\phi$ preserves
    structure, thus we need to show $(1), (2),$ and $(3)$. Starting with
    number one:
    \[ \text{\bf (1)} \hspace{0.3cm} \phi(L+V) = \phi(L) \cup \phi(V) \]
    In the following let $L,V \in \fps{B}$ arbitrary but fixed. Then
    \begin{align*}
      \phi(L+V) &= \{ w \in \Sigma^* : (L,w) + (V,w) = 1 \} \\
        &= \{ w \in \Sigma^* : (L,w) = 1 \lor (V,w) = 1 \} \\
        &= \{ w \in \Sigma^* : w \in \{ x \in \Sigma^* : (L,x) = 1 \}
        \lor w \in \{ x \in \Sigma^* :(V,x) = 1 \} \} \\
        &= \{ w \in \Sigma^* : w \in \phi(L) \lor w \in \phi(V) \} 
        = \phi(L) \cup \phi(V).
    \end{align*}
    Continuing with two:
    \[ 
      \text{\bf (2)} \hspace{0.3cm} 
      \phi(L\cdot V) = \phi(L) \cdot \phi(V) 
    \]
    In the following let $L,V \in \fps{B}$ arbitrary but fixed. Then
    \begin{align*}
      \phi(L\cdot V) &= \{ uv \in \Sigma^* : (L,u) \cdot (V,v) = 1\}\\
        &= \{ uv \in \Sigma^* : (L,u) = 1 \land (V,v) = 1\} \\
        &= \{ uv \in \Sigma^* : u \in \{ w \in \Sigma^* : (L,w) = 1\}
          \land v \in \{ w \in \Sigma^* : (V,w) = 1\}
        \} \\
        &= \{ uv \in \Sigma^* : u \in \phi(L)
          \land v \in \phi(V)
        \} = \phi(L) \cdot \phi(V).
    \end{align*}
    Last but not least, three:
    \[ 
      \text{\bf (3)} \hspace{0.3cm} 
      \phi(L^*) = \phi(L)^*
    \]
    In the following let $L \in \fps{B}$ arbitrary but fixed. Then
    \begin{align*}
      \phi(L^*) = \phi(\sum_{i\in \mathbb{N}_0} L^i) 
      \stackrel{\text{\bf (1)}}{=} \sum_{i\in \mathbb{N}_0} \phi(L^i)
      \stackrel{\text{\bf (2)}}{=} \sum_{i\in \mathbb{N}_0} \phi(L)^i
      = \phi(L)^*
    \end{align*}
    Thus $\phi$ is an isomorphism and our claim follows.
\end{proof}
\\\\
Now we have shown that both definition are equivalent, but why should
one use the less natural formal power series definition? The anwser
to this, as so often in mathematics, is {\bf abstraction}. Finite automata
are accepting formal languages, thus we can say they generate a formal
language. When abstracting from finite automata's to weighted automata's
which have a weight or cost between each transition, the generated
object is not longer a language, but a map that assigns a weight or
cost to each word. Generally, this weight or cost can be seen as a 
semiring (a Ring that does not have necessarily an additive inverse),
which are objects that behave like numbers, we will denote an arbitrary
semiring with $\mathbb{S}$. Those maps, that get accepted by an weighted
automata, have the signature $L: \Sigma^* \to \mathbb{S}$ and are 
indeed a formal power series over $\mathbb{S}$, consequently $L \in \fps{S}$.
To Summarize, finite automata generate formal power series over $\mathbb{B}$
and weighted automata generate formal power series over $\mathbb{S}$. So 
formal power series are in fact an abstraction of formal languages.
\section{Kleene's theorem (Power series, Matrix Automata)}
In this chapter we will examine, what kind of languages get accepted 
by an finite automata. In order to do that, we will prove that all
languages that get accepted by an finite automata are rational and every
rational language can be generated by an finite automata (kleene's theorem).
A rational language is defined inductively.
\begin{definition}
  Let $\Sigma$ be an fixed alphabet, then a rational language 
  is defined as follows:
  \begin{enumerate}
    \item A language $L = a \in \fps{B}$ with a single character 
      $a \in \Sigma$ is said to be rational.
    \item The empty word $\epsilon$ is rational.
    \item The sum of two rational Languages $L_1, L_2 \in \fps{B}$
      are again rational $L = L_1 + L_2$.
    \item The product of two rational Languages $L_1, L_2 \in \fps{B}$
      are again rational $L = L_1 \cdot L_2$.
    \item The kleene star of a rational language $L \in \fps{B}$ 
        is again rational $V = L^*$.
  \end{enumerate}
\end{definition}
So we start with just one character and then building with our operations
step by step  a language. To show kleene's theorem we have to prove two
direction. First we will prove that finite automata only accept rational 
languages. Afterwards, we will prove that every rational language is 
accepted by an finite automata. In an effort to do that, we need to 
understand certain properties of the Matrix encoding of finite automatas.
Raising the power of two of a Matrix is defined by:\\
Let $A \in \fps{B}^{n\times n}$ be an arbitrary matrix, then
\[ 
  (A^2)_{ij} = \sum_{k=1}^{n} A_{ik}A_{kj}.
\]
As mentioned before, this matrix is a transistion function of an finite
automata, where we can look up, what character carry one state in another.
In $A_{ij}$ is the character that carry $i$ to $j$, thus multiplying
$A$ by itself, gives us a new Matrix with all words with length two,
given as input and $i$ is the starting node would
carry $i$ to $j$ in two steps. In the formula above we see that we
are trying to go throug all nodes $k \in \{1, 2, 3, \dots n\}$
to end up in $j$ and then uniting them together with addition.\\
Generally, we can have a matrix $A \in \fps{B}^{n \times n}$
risen to the power of $n \in \mathbb{N}$:
\[
  (A^n)_{ij} = \sum_{k_1=1}^{n}
    \sum_{k_2=1}^{n}\dots \sum_{k_{n-1}=1}^n 
    a_{ik_1}a_{k_1k_2}a_{k_2k_3}\dots a_{k_{n-1}j} \\
\]
This is not beatiful at all, but the idea stays exactly the same: Now
we have in $(A^n)_{ij}$ all words that carry $i$ to $j$ in n steps,
brute forcing throug all possible paths. Now to get all possible paths
of arbitrary length we just add all together:
\[
  E_n + A + A^2 \dots = \sum_{i \in \mathbb{N}} A^i = A^*
\]
Yielding, that in $(A^*)_{ij}$ are all possible words, carrying
the automata from $i$ to $j$. If $i$ is the starting node and $j$
the accepting node, the accepting language of the automata is just
$(A^*)_{ij}$. That means, $A^*$ stores all accepting languages of 
an automata, with transition matrix $A$.
With those important properties, we can now move on to the first prove.
\begin{proof}
  \begin{center}
    {\bf claim: All finite automata only accept rational languages.}
  \end{center}
  Let $A \in \fps{B}^{n \times n}$ be a transition function of an
  arbitrary finite automata. $A$ is rational, because the transition
  function consists only out of characters. Since $A^*$ stores every accepting
  language of that finite autotamata, for every possible initial
  state and accepting state, we just need to prove that all entries
  in $A^*$ are rational languages.
  We will prove this statement via induction over the number of 
  states, namely the dimension of the square matrix $n \in \mathbb{N}$.\\
  {\bf Base case $n=1$}:\\
  Thus
  \[
    A \in \fps{B}^{1 \times 1} = \fps{B}.
  \]
  We know that $A \in \fps{B}$ is rational, but that also means
  that $A^*$ is per definition rational.

\end{proof}

\section{Maximum Gap Lemma}
\section{Not all Formal languages are regular}
\section{Kleene Schützenberg theorem and weigthed automatas}
\section{Conclusion}

\subsection{Citation} 
\label{sec:cit}

The citation method follows the author-year system. Place reference is in the text, footnotes should only be used for explanations and comments. The following notes are taken from the \emph{language} bibliography template from \url{ron.artstein.org}:\newline

\noindent
The \emph{Language} style sheet makes a distinction between two kinds of in-text citations: citing a work and citing an author.
\begin{itemize}
\item Citing a work:
  \begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parsep}{0pt}
  \item Two authors are joined by an ampersand (\&).
  \item More than two authors are abbreviated with \emph{et al.}
  \item No parentheses are placed around the year (though parentheses
    may contain the whole citation). 
  \end{itemize}
\item Citing an author:
  \begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parsep}{0pt}
  \item Two authors are joined by \emph{and}.
  \item More than two authors are abbreviated with \emph{and colleagues}.
  \item The year is surrounded by parentheses (with page numbers, if
    present).
  \end{itemize} 
\end{itemize}
To provide for both kinds of citations, \verb+language.bst+ capitalizes on the fact that \verb+natbib+ citation commands come in
two flavors. In a typical style compatible with \verb+natbib+, ordinary commands such as \verb+\citet+ and \verb+\citep+ produce short
citations abbreviated with \emph{et al.}, whereas starred commands such as \verb+\citet*+ and \verb+\citep*+ produce a citation with a
full author list. Since \emph{Language} does not require citations with full authors, the style \verb+language.bst+ repurposes the starred commands to be used for citing the author. The following table shows how the \verb+natbib+ citation commands work with \verb+language.bst+.
\begin{center}
  \begin{tabular}{lll}
    \toprule
    Command & Two authors & More than two authors \\
    \midrule
    \verb+\citet+ & \citet{hale} & \citet{sprouse} \\
    \verb+\citet*+ & \citet*{hale} & \citet*{sprouse} \\
    \addlinespace
    \verb+\citep+ & \citep{hale} & \citep{sprouse} \\
    \verb+\citep*+ & \citep*{hale} & \citep*{sprouse} \\
    \addlinespace
    \verb+\citealt+ & \citealt{hale} & \citealt{sprouse} \\
    \verb+\citealt*+ & \citealt*{hale} & \citealt*{sprouse} \\
    \addlinespace
    \verb+\citealp+ & \citealp{hale} & \citealp{sprouse} \\
    \verb+\citealp*+ & \citealp*{hale} & \citealp*{sprouse} \\
    \addlinespace
    \verb+\citeauthor+ & \citeauthor{hale} & \citeauthor{sprouse} \\
    \verb+\citeauthor*+ & \citeauthor*{hale} & \citeauthor*{sprouse} \\
    \verb+\citefullauthor+ & \citefullauthor{hale} & \citefullauthor{sprouse} \\
    \bottomrule
  \end{tabular}
\end{center}
Authors of \emph{Language} articles would typically use \verb+\citet*+, \verb+\citep+, \verb+\citealt+ and \verb+\citeauthor*+, though they
could use any of the above commands. There is no command for giving a full list of authors.

\subsection{Bibliography}
The bibliography of this template includes the references of the \emph{language} stylesheet as a sample bibliography.

\pagebreak

\microtypesetup{protrusion=false}
\listoffigures{}
\listoftables{}
\microtypesetup{protrusion=true}

\clearpage
\printglossaries
\pagebreak
\addcontentsline{toc}{section}{Bibliography}
\pagestyle{fancy}
\bibliographystyle{language-dt} 
\bibliography{bibliography} 
\nocite{*}
\end{document}